---
title: "Assignment 2: Coding Basics"
author: "Jiyoung Park"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics Day 1

1.  Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r}
#1. 
sequence_by_four <- seq(1,100,4)
sequence_by_four 

#2. 
result_mean <- mean(sequence_by_four)
result_mean
result_median <-median(sequence_by_four)
result_median

#3. 
if(result_mean > result_median){ #comparing two values 
  print (result_mean) #print mean if mean is greater 
}else {
  print (result_median) #print median elsewise 
}

```

## Basics Day 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r}
student_names <- c("Jin","RM", "Jimin","Jungkook") #character vector
student_names
test_score <- c(85,95,70,40) #numeric vector
test_score
if_passed <- c(TRUE,TRUE,TRUE,FALSE) #logical vector
if_passed

test_result <-data.frame(student_names,test_score, if_passed)
names(test_result) <- c("Name","Score","Passed"); View(test_result)


```

9.  QUESTION: How is this data frame different from a matrix?

> Answer: matrix can only have the same data type while data frame can have differnt types of data 

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

11. Apply your function to the vector with test scores that you created in number 5.

```{r}

ifelse(test_score >= 50,TRUE,FALSE)

```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: ifelse worked because vector has more than one length and ifelse creates a "loop" which allows it to go over all the test scores in the vector. 
