---
title: "Assignment 5: Data Visualization"
author: "Jiyoung Park "
output: pdf_document
geometry: margin=2.54cm
---

```{r setup, include=FALSE}
#This code chunk will tidy your knit PDF files, wrapping long code lines
#For it to work, the "formatR" package needs to be installed

#install.packages('formatR')
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Visualization 

## Directions
1. Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.

The completed exercise is due on Friday, Oct 21th @ 5:00pm.



## Set up your session 
1. Set up your session. Verify your working directory and load the tidyverse, 
lubridate, & cowplot packages. Upload the NTL-LTER processed data files for  
nutrients and chemistry/physics for Peter and Paul Lakes (use the tidy 
[`NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv`] version) and 
the processed data file for the Niwot Ridge litter dataset (use the
[`NEON_NIWO_Litter_mass_trap_Processed.csv`] version). 

2. Make sure R is reading dates as date format; if not change the format to date.

```{r setup, warning=FALSE}
#1 
setwd("~/EDA-Fall2022")
require("knitr")
opts_knit$set(root.dir = "~/EDA-Fall2022")
getwd()
library(tidyverse)
library(lubridate)
#install.packages("cowplot")
library(cowplot)
NTL_LTER <- read.csv("./Data/Processed/NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv", stringsAsFactors = TRUE)
NEON_NIWO <- read.csv("./Data/Processed/NEON_NIWO_Litter_mass_trap_Processed.csv", stringsAsFactors = TRUE)
#2 
NEON_NIWO$collectDate <- as.Date(NEON_NIWO$collectDate, format = "%Y-%m-%d")
NTL_LTER$sampledate <- as.Date(NTL_LTER$sampledate, format = "%Y-%m-%d")
```
## Define your theme
3. Build a theme and set it as your default theme.

```{r}
#3
mytheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom") 

```

## Create graphs
For numbers 4-7, create ggplot graphs and adjust aesthetics to follow best 
practices for data visualization. Ensure your theme, color palettes, axes, 
and additional aesthetics are edited accordingly.

4. [NTL-LTER] Plot total phosphorus (`tp_ug`) by phosphate (`po4`), with 
separate aesthetics for Peter and Paul lakes. Add a line of best fit and 
color it black. Adjust your axes to hide extreme values (hint: change the 
limits using `xlim()` and/or `ylim()`).

```{r , warning=FALSE}
#4
ggplot(NTL_LTER, aes(x=po4, y=tp_ug))+
 geom_point(aes(color= lakename))+
 geom_smooth(method =lm, color = "black", se=FALSE) +
  mytheme+
 xlim(0,50) +ylim(0,150)+
  labs(title ="Phosphorus vs. Phosphate", x = "Phosphate", y = "Phosphorus")

#zooming in more 
ggplot(NTL_LTER, aes(x=po4, y=tp_ug))+
 geom_point(aes(color= lakename))+
 geom_smooth(method =lm, color = "black", se=FALSE) +
  mytheme+
 xlim(0,10) +ylim(0,50)+
    labs(title ="Phosphorus vs. Phosphate", x = "Phosphate", y = "Phosphorus")

```


5. [NTL-LTER] Make three separate boxplots of (a) temperature, (b) TP, and 
(c) TN, with month as the x axis and lake as a color aesthetic. Then, create 
a cowplot that combines the three graphs. Make sure that only one legend is 
present and that graph axes are aligned. 

Tip: R has a build in variable called `month.abb` that returns a list of months;
see <https://r-lang.com/month-abb-in-r-with-example>

```{r setup, warning=FALSE}
#5
abbmonth <-month.abb[NTL_LTER$month]

temp <- ggplot(NTL_LTER, aes(x= abbmonth, y= temperature_C)) +
  geom_boxplot(aes(color= lakename))+
   mytheme+
  theme(legend.position = "none")+
  labs(title ="Temperature by months", x= "months", y= "Temperature (C)")

TP <- ggplot(NTL_LTER, aes(x= abbmonth, y= tp_ug))+
  geom_boxplot(aes(color= lakename))+
   mytheme+
  theme(legend.position = "none")+
  labs(title ="TP by months", x= "months", y= "TP")
  
TN <- ggplot(NTL_LTER, aes(x= abbmonth, y= tn_ug))+
  geom_boxplot(aes(color= lakename))+
   mytheme+
  theme(legend.position = "none")+
  labs(title ="TN by months", x= "months", y= "TN")   

temp_legend <- ggplot(NTL_LTER, aes(x= abbmonth, y= temperature_C)) +
  geom_boxplot(aes(color= lakename))
legend <- get_legend(temp_legend)

plot_grid(temp,TP,TN, legend, align = 'h')

```

Question: What do you observe about the variables of interest over seasons 
and between lakes? 

> Answer: high TP and TN concentration with months that are high in temperatures. No measures found in months with cold temperatures(Feb, Nov, Oct)

6. [Niwot Ridge] Plot a subset of the litter dataset by displaying only the 
"Needles" functional group. Plot the dry mass of needle litter by date and 
separate by NLCD class with a color aesthetic. (no need to adjust the name 
of each land use)

7. [Niwot Ridge] Now, plot the same plot but with NLCD classes separated into 
three facets rather than separated by color. 

```{r setup, warning=FALSE}
#6
NEON_NIWO_needles <- filter(NEON_NIWO,functionalGroup == "Needles")
ggplot(NEON_NIWO_needles,aes(x= collectDate, y=dryMass)) +
  geom_point(aes(color=nlcdClass)) +
  labs(title ="Needles Litter by date", x = "Date", y ="Dry Mass")


#7
ggplot(NEON_NIWO_needles,aes(x= collectDate, y=dryMass)) +
  geom_point() +
  facet_wrap(vars(nlcdClass)) +
  labs(title ="Needles Litter by date", x = "Date", y ="Dry Mass")


```
Question: Which of these plots (6 vs. 7) do you think is more effective, and why? 

> Answer: i believe plot 7 is more effective at comparing than plot 6 because i can see the trends for drymass for three different ncld class more clearly in plot 7. 